---
title: "Analyse Endnote Library"
author: "Michael Rustler"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Analyse Endnote Library}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


# Import Endote XML As Data Frame

```{r echo=TRUE}
endnote_xml <- system.file("extdata/KWB_documents.xml", 
                           package = "kwb.endnote")

references_df <- kwb.endnote::create_df_from_endnote_xml(endnote_xml)

n_publications <- length(unique(references_df$record_id))

```

# Analysis of Publications

The imported Endnote XML library `r basename(endnote_xml)` contains 
`r n_publications` which are analysed in detail in the following. 

## By Type


```{r echo=TRUE}
library(dplyr)
refs_by_type <- kwb.endnote::get_reference_type_names(endnote_xml) %>%
  dplyr::count(.data$ref_type_name) %>%
  dplyr::arrange(dplyr::desc(.data$n))

knitr::kable(refs_by_type)
```


## Has an Abstract?

```{r echo=TRUE}
refs_with_abstract <- references_df %>% 
  dplyr::filter(.data$key1 == "abstract") %>%
  dplyr::count(.data$record_id, .data$ref_type_name) 

n_pubs_with_abstract <- nrow(refs_with_abstract) 
### Percent of Publications with Abstracts
percent_pubs_with_abstracts <- round(100*nrow(refs_with_abstract)/n_publications,1) 

```

`r percent_pubs_with_abstracts` percent (i.e. 
`r sprintf("%d / %d", n_pubs_with_abstract, n_publications)` 
publications) have an abstract. However there are some line breaks for 


## By Accessibility Level 

Is the accessibility level for the referenc defined in the fiel"custom3" 
(confidential or not?)

```{r echo=TRUE}
refs_with_accessability_level <- references_df %>% 
  dplyr::filter(.data$key1 == "custom3") %>%
  dplyr::count(.data$value) 

n_pubs_with_accessability_level <- nrow(refs_with_accessability_level) 

percent_with_accessibility_level <- round(100*n_pubs_with_accessability_level/n_publications, 1)
knitr::kable(refs_with_accessability_level)

is_confidential <- stringr::str_detect(refs_with_accessability_level$value, 
                         stringr::regex("conf|PU|intern|private",
                                        ignore_case = TRUE))

n_pubs_confidential <-sum(refs_with_accessability_level$n[is_confidential])

### in percent
percent_confidential <- 100*sum(refs_with_accessability_level$n[is_confidential])/n_publications 
```

Only for `r n_pubs_with_accessability_level` (i.e. `r percent_with_accessibility_level` percent) 
meta information on the accessibility level is explicitly defined, out of which 
`r n_pubs_confidential` are defined as confidential.


## By Project 

Group references by "custom2" (i.e. "project names")

```{r echo=TRUE}
refs_with_project <- references_df %>% 
  dplyr::filter(.data$key1 == "custom2") 


n_pubs_with_project_meta <- nrow(refs_with_project)

### in percent
percent_pubs_with_project_meta <- round(100*n_pubs_with_project_meta/n_publications,1)


unique_project_names <- refs_with_project  %>% 
  dplyr::count(.data$value) 

knitr::kable(unique_project_names)

```

In total `r percent_pubs_with_project_meta` percent (i.e. 
`r sprintf("%d / %s", n_pubs_with_project_meta, n_publications)` publications) 
contain meta-information on the project. However, the spelling of project names
does not follow a controlled vocabulary yet as can be seen in the table above. 

Thus it is recommended to establish a controlled vocabulary for [project acroyms](https://kwb-r.github.io/fakin.doc/best-practices.html#acronyms) 
according to the best-practices defined in the FAKIN project (see [here](https://kwb-r.github.io/fakin.doc/best-practices.html#acronyms))

## By Author

## Lastname/Firstname

Quite messy. Names should be entered in Endnote in the following format:

**Lastname, Firstname**


```{r echo=TRUE, fig.height=10, fig.width=8}

refs_by_author_lastfirst <- references_df %>%
  dplyr::filter(.data$key2 == "authors") %>%
  dplyr::count(.data$value)  %>%
  dplyr::arrange(dplyr::desc(.data$n))

kwb.endnote::plot_pubs_by_author(refs_by_author_lastfirst[1:50, ])

```

## Lastname

For now data-munging is needed to just use the authors` lastname for aggregating 
the data

```{r echo=TRUE, fig.height=10, fig.width=8}
refs_by_author_last <- references_df %>%
  dplyr::filter(.data$key2 == "authors") %>%
  dplyr::mutate(value = stringr::str_remove_all(.data$value, ",.*")) %>%
  dplyr::mutate(value = stringr::str_remove_all(.data$value, "^\\w+\\.?\\s+")) %>%
  dplyr::count(.data$value)  %>%
  dplyr::arrange(dplyr::desc(.data$n))

kwb.endnote::plot_pubs_by_author(refs_by_author_last[1:50, ])

#plotly::ggplotly(g)

```
