---
title: "Analyse Endnote Library"
author: "Michael Rustler"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Analyse Endnote Library}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


# Import Endnote XML As Data Frame

```{r echo=TRUE}
library(kwb.endnote)

endnote_xml <- kwb.endnote::default_xml()

references_df <- kwb.endnote::create_df_from_endnote_xml(endnote_xml)

n_publications <- length(unique(references_df$record_id))
```

# Analysis of Publications

The imported Endnote XML library `r basename(endnote_xml)` contains 
`r n_publications` which are analysed in detail in the following. 

## By Type


```{r echo=TRUE}
refs_by_type <- kwb.endnote::get_reference_type_names(endnote_xml) %>%
  dplyr::count(.data$ref_type_name) %>%
  dplyr::arrange(dplyr::desc(.data$n))

knitr::kable(refs_by_type)
```

## By Year

```{r echo=TRUE}
endnote_list <- kwb.endnote::create_endnote_list(endnote_xml)
refs_df <- kwb.endnote::create_references_df(endnote_list)

refs_by_year <- refs_df %>%
dplyr::count(ref_type_name, year)
  
knitr::kable(refs_by_year)
```  
  
```{r echo=TRUE, fig.width=8, fig.height=9}

kwb.endnote::plot_pubs_by_year(refs_df)

```

## By Keywords

```{r echo=TRUE}

  keywords_df <- kwb.endnote::create_keywords_df(references_df) 
 
  keywords_df %>% 
  dplyr::filter(.data$n > 1) %>% 
  knitr::kable()
```

Create a word cloud for the keywords

```{r echo=TRUE, fig.width=7, fig.height=7}

kwb.endnote::plot_wordcloud_keywords(keywords_df)
```

## Which fields are filled?

## Ordered Alphabetically
```{r echo=TRUE}
references_df %>%  
dplyr::count(key1, key2, key3, key4) %>%  
knitr::kable()
```

## Ordered By Frequency
```{r echo=TRUE}
references_df %>%  
dplyr::count(key1, key2, key3, key4) %>%  
dplyr::arrange(dplyr::desc(n)) %>% 
knitr::kable()
```

## Ordered By Ref Type Name and Frequency

```{r echo=TRUE}
references_df %>%  
dplyr::count(ref_type_name, key1, key2, key3, key4) %>%  
dplyr::arrange(ref_type_name, dplyr::desc(n)) %>% 
knitr::kable()
```


## Has an Abstract?

```{r echo=TRUE}
refs_with_abstract <- references_df %>% 
  dplyr::filter(.data$key1 == "abstract") %>%
  dplyr::count(.data$record_id, .data$rec_number, .data$ref_type_name) 

n_pubs_with_abstract <- nrow(refs_with_abstract) 
### Percent of Publications with Abstracts
percent_pubs_with_abstracts <- round(100*nrow(refs_with_abstract)/n_publications,1) 

```

`r percent_pubs_with_abstracts` percent (i.e. 
`r sprintf("%d / %d", n_pubs_with_abstract, n_publications)` 
publications) have an abstract. 

```{r echo=TRUE}
refs_with_abstract_and_linebreaks <- refs_with_abstract %>%  
  dplyr::filter(.data$n > 1) %>% 
  dplyr::arrange(dplyr::desc(.data$n))

n_pubs_with_abstract_and_linebreaks <- nrow(refs_with_abstract_and_linebreaks)

knitr::kable(refs_with_abstract_and_linebreaks)

```

However there are some line breaks for these abstracts for in total 
`r n_pubs_with_abstract_and_linebreaks` publications as shown in the table above. 

These need to be corrected manually!


## By Accessibility Level 

Is the accessibility level for the referenc defined in the fiel"custom3" 
(confidential or not?)

```{r echo=TRUE}
refs_with_accessability_level <- references_df %>% 
  dplyr::filter(.data$key1 == "custom3") %>%
  dplyr::count(.data$value) 

n_pubs_with_accessability_level <- sum(refs_with_accessability_level$n) 

percent_with_accessibility_level <- round(100*n_pubs_with_accessability_level/n_publications, 1)
knitr::kable(refs_with_accessability_level)


is_confidential <- grepl(paste(c("conf", "PU", "intern", "private"), collapse="|"), 
                        refs_with_accessability_level$value)

n_pubs_confidential <-sum(refs_with_accessability_level$n[is_confidential])

### in percent
percent_confidential <- 100*sum(refs_with_accessability_level$n[is_confidential])/n_pubs_with_accessability_level
```

Only for `r n_pubs_with_accessability_level` (i.e. `r percent_with_accessibility_level` percent) 
meta information on the accessibility level is explicitly defined, out of which 
`r n_pubs_confidential` are defined as confidential (i.e. 
`r round(percent_confidential,1)` of the publications with metadata on the 
accessibility level).

**Recommendation:**

Fill the field **custom3** for all publications with either: 

- **public** or 

- **confidential** 




## By Project 

Group references by "custom2" (i.e. "project names")

```{r echo=TRUE}
refs_with_project <- references_df %>% 
  dplyr::filter(.data$key1 == "custom2") 


n_pubs_with_project_meta <- nrow(refs_with_project)

### in percent
percent_pubs_with_project_meta <- round(100*n_pubs_with_project_meta/n_publications,1)


unique_project_names <- refs_with_project  %>% 
  dplyr::count(.data$value) 

knitr::kable(unique_project_names)

```

In total `r percent_pubs_with_project_meta` percent (i.e. 
`r sprintf("%d / %s", n_pubs_with_project_meta, n_publications)` publications) 
contain meta-information on the project. However, the spelling of project names
does not follow a controlled vocabulary yet as can be seen in the table above. 

Thus it is recommended to establish a controlled vocabulary for [project identifiers](https://kwb-r.github.io/fakin.doc/qms.html#project-identifier) 
according to the best-practices defined in the FAKIN project (see 
[here](https://kwb-r.github.io/fakin.doc/best-practices.html#acronyms)).

## By Author

## Lastname/Firstname



```{r echo=TRUE, fig.height=10, fig.width=8}

refs_by_author_lastfirst <- references_df %>%
  dplyr::filter(.data$key2 == "authors") %>%
  dplyr::count(.data$value)  %>%
  dplyr::arrange(dplyr::desc(.data$n))

kwb.endnote::plot_pubs_by_author(refs_by_author_lastfirst[1:50, ])

```

Quite messy the figure above. Thus names should be entered in Endnote in the 
following format:

**Lastname, Firstname**

Also note the **,** also is an author! This is due to entering a wrong semicolon 
in the Endnote database and needs to be fixed here for at least the following 
entries:


```{r echo=TRUE}

pubs_with_semicolon_as_author <- references_df %>%
  dplyr::filter(.data$key2 == "authors",
                .data$value == ",") %>% 
  dplyr::count(.data$rec_number) %>% 
  dplyr::arrange(dplyr::desc(.data$n)) 


n_pubs_with_semicolon_as_author <- nrow(pubs_with_semicolon_as_author)

knitr::kable(pubs_with_semicolon_as_author)
```

In total `r n_pubs_with_semicolon_as_author` publications contain one or more 
**semicolon** authors. This needs to be fixed manually in the Endnote DB. 

## Lastname

For now data-munging is needed to just use the authors` lastname for aggregating 
the data

```{r echo=TRUE, fig.height=10, fig.width=8}
refs_by_author_last <- references_df %>%
  dplyr::filter(.data$key2 == "authors") %>%
  dplyr::mutate(value = stringr::str_remove_all(.data$value, ",.*")) %>%
  dplyr::mutate(value = stringr::str_remove_all(.data$value, "^\\w+\\.?\\s+")) %>%
  dplyr::count(.data$value)  %>%
  dplyr::arrange(dplyr::desc(.data$n))

kwb.endnote::plot_pubs_by_author(refs_by_author_last[1:50, ])


```

## Journal Articles

### By Journal

Journal names need to be harmonised. Also needs to be checked what the differences
between **full-title** and **secondary-title** of journal are.

#### Journal full-title

Defined in: *$`record`$periodical$`full-title`$`style`[[1]]*

```{r echo=TRUE}

pubs_in_journals <- references_df %>%  
  dplyr::filter(.data$ref_type_name == "Journal Article")

n_pubs_in_journals <- length(unique(pubs_in_journals$record_id))

  
journal_title_full <- pubs_in_journals %>%  
  dplyr::filter(.data$key2 %in% c("full-title")) %>% 
  dplyr::mutate("source_field" = sprintf("%s_%s",.data$key1, 
                                           .data$key2))  %>% 
  dplyr::count(.data$value, .data$source_field) 

knitr::kable(journal_title_full)


```

#### Journal secondary-title 

Defined in: *$`record`$titles$`secondary-title`$`style`[[1]]*

```{r echo=TRUE}
journal_title_secondary <- pubs_in_journals %>%  
  dplyr::filter(.data$key2 %in% c("secondary-title")) %>% 
  dplyr::mutate("source_field" = sprintf("%s_%s",.data$key1, 
                                           .data$key2))  %>% 
  dplyr::count(.data$value, .data$source_field)

knitr::kable(journal_title_secondary)


```

#### Journal DOI

All `r n_pubs_in_journals` journal articles should have an DOI. Let`s check 
for how many we have this metadata:


```{r echo=TRUE}
has_doi <- stringr::str_detect(pubs_in_journals$value, "doi")
pubs_in_journals_with_doi <- pubs_in_journals[has_doi, ]

n_pubs_in_journals_with_doi <- nrow(pubs_in_journals_with_doi)

knitr::kable(pubs_in_journals_with_doi)
```

Only `r n_pubs_in_journals_with_doi` journal publications have a DOI which is 
either encoded in the field **electronic-resource-num** and **urls**.

All DOIs should be entered in the field **electronic-resource-num** in the 
following format, i.e.:

**https://doi.org/10.6084/m9.figshare.828487**


# Export References To XLSX

Write references dataframe to xlsx with one sheet for each **publication type**
to `references.xlsx`

```{r echo=TRUE, eval=FALSE}
kwb.endnote::write_references_df_to_xlsx(endnote_list)
```

And also writting a **cleaned** version after automatically running 
`kwb.endnote::clean_references_df()` to `references_cleaned.xlsx`.

```{r echo=TRUE, eval=FALSE}
kwb.endnote::write_clean_references_df_to_xlsx(endnote_list)
```

# Check Endnote References

```{r echo=TRUE, eval=TRUE}
problematic_entries <- kwb.endnote::check_problematic_entries(endnote_list)
knitr::kable(problematic_entries[1:100,])
```

